---
title: "Rasheed Wallace Waffles"
author: "Chris Gallo"
date: 2020-07-17
categories: ["R"]
slug: "sheed-wallace-waffles"
type: "post"
tags: ["R Markdown", "Fungo Friday"]
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r table, echo=FALSE}
library(ggplot2)
library(waffle)
library(hrbrthemes)
library(htmltools)

sheed.college <- c(`3PTers (0.3%)` = 1, `2PTers (83.7%)` = 84, `Free Throws (16.0%)` = 15)

waffle(
  sheed.college, rows = 5, size = 4,
  colors = c("#969696", "#56a0d3", "#13294b"),
  title = "Rasheed Wallace: College Points Distribution (69 Games)"
) -> College

sheed.nba <- c(`3PTers (20.5%)` = 21, `2PTers (65.5%)` = 65, `Free Throws (14.2%)` = 14)

waffle(
  sheed.nba, rows = 5, size = 4,
  colors = c("#ed174c", "#0067b1", "#002d62"),
  title = "Rasheed Wallace: NBA Points Distribution (1109 Games)"
) -> NBA


iron(College, NBA)
```
---

Rasheed Wallace is one of my favorite basketball players. This uses a waffle chart to display his points distribution in both college and the NBA. 

Wallace attempted only four three-point shots in his collegiate career. This means only 0.67 percent of his shot attempts at North Carolina were behind the three-point line. 

He played 18 years in the NBA and attempted 3,228 three-point shots. 23.9 percent of his shots came from three-point land as a pro, and he scored ~21 percent of his points in the NBA via the three-point shot.

Built using the [waffle package](https://github.com/hrbrmstr/waffle). 

Data via [UNC Record Book](https://goheels.com/documents/2019/10/30/2019_19_UNC_Men_s_Basketball_Record_Book.pdf) and [Basketball Reference](https://www.basketball-reference.com/players/w/wallara01.html).

Code is below . . .

```
library(ggplot2)
library(waffle)
library(hrbrthemes)
library(htmltools)

sheed.college <- c(`3PTers (0.3%)` = 1, `2PTers (83.7%)` = 84, `Free Throws (16.0%)` = 15)

waffle(
  sheed.college, rows = 5, size = 4,
  colors = c("#969696", "#56a0d3", "#13294b"),
  title = "Rasheed Wallace: College Points Distribution (69 Games)"
) -> College

sheed.nba <- c(`3PTers (20.5%)` = 21, `2PTers (65.5%)` = 65, `Free Throws (14.2%)` = 14)

waffle(
  sheed.nba, rows = 5, size = 4,
  colors = c("#ed174c", "#0067b1", "#002d62"),
  title = "Rasheed Wallace: NBA Points Distribution (1109 Games)"
) -> NBA


iron(College, NBA)
```
