---
title: "Bells are not Dumb"
author: "Chris Gallo"
date: 2020-08-07
categories: ["R"]
slug: "dumb-bell"
type: "post"
tags: ["R Markdown", "Fungo Friday"]
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.width=12, fig.height=8)
```

```{r, echo=FALSE, message=FALSE}
library(tidyverse)
library(dplyr)
library(ggalt)

# fix and manipulate data
d <- tibble::tribble(
    ~player, ~college, ~nba,
  "Rasheed Wallace",  0.01, 0.24, 
  "Jerry Stackhouse", 0.15, 0.24, 
  "Vince Carter", 0.31, 0.29, 
  "Antawn Jamison", 0.02, 0.20 
  )

d %>% 
  mutate(diff = nba - college) -> d 

d$player <- factor(d$player, levels = rev(d$player))

d %>%
  ggplot() +
  geom_dumbbell(aes(y = player,
      x = college, xend = nba),
    colour = "grey60", size = 5,
    colour_x = "#56a0d3", colour_xend = "#395B74") +
  theme_minimal() +
  geom_text(data = filter(d, player == "Rasheed Wallace"),
    aes(x = college, y = player),
    label = "College", fontface = "bold",
    color = "#56a0d3",
    vjust = -1) +
  geom_text(data = filter(d, player == "Rasheed Wallace"),
    aes(x = nba, y = player),
    label = "NBA", fontface = "bold",
    color = "#395B74",
    vjust = -1) +
  labs(x = "Three-Point Attempt Rate (% of shots from 3)",
    y = "Player",
    title = "Three-Point Attempt Rates in Collegiate and Pro Careers", caption = "Source: UNC Record Book and Pro-Basketball Reference") +
  theme(title = element_text(size = rel(1.4))) +
  scale_x_continuous(labels = scales::percent,
    breaks = c(0, 0.10, 0.20, 0.30, .40)) +
  coord_cartesian(xlim = c(0, 0.40)) +
  geom_text(aes(y = player, label = diff),
    x = 0.4, hjust = 0.25) +
  annotate(x = 0.4, y = "Rasheed Wallace", label = "Delta",
    geom = "text", vjust = -2,
    fontface = "bold",
    hjust = 1)

```
---

This is a riff off the [Rasheed Wallace Waffles](https://fungo.hotdogsandeggs.com/sheed-wallace-waffles/) post. It looks at college and pro three-point rates for four Tar Heels that played at least 16 seasons in the NBA using a dumbbell plot to show the change. 

It's incredible how much the game changed, and Vince Carter's rate actually went down. Carter played 22 years in the league. About 31 percent of Carter's shot attempts were from thee-point land as a collegiate player, and about 29 percent as a pro. 

Carter's three-point percentage was 36.8 in college, and a 37.1 in the NBA. Consistency. 

The data was pulled from [basketball-reference.com](https://www.basketball-reference.com/) and the [UNC Record Book](https://goheels.com/documents/2019/10/30/2019_19_UNC_Men_s_Basketball_Record_Book.pdf). This plot uses the [ggalt package](https://cran.r-project.org/web/packages/ggalt/index.html). The [tutorial here](https://data-se.netlify.app/2018/05/23/playing-around-with-dumbbell-plots/)) was helpful. 

Code is below . . .

```
library(tidyverse)
library(dplyr)
library(ggalt)

# fix and manipulate data
d <- tibble::tribble(
    ~player, ~college, ~nba,
  "Rasheed Wallace",  0.01, 0.24, 
  "Jerry Stackhouse", 0.15, 0.24, 
  "Vince Carter", 0.31, 0.29, 
  "Antawn Jamison", 0.02, 0.20 
  )

d %>% 
  mutate(diff = nba - college) -> d 

d$player <- factor(d$player, levels = rev(d$player))

d %>%
  ggplot() +
  geom_dumbbell(aes(y = player,
      x = college, xend = nba),
    colour = "grey60", size = 5,
    colour_x = "#56a0d3", colour_xend = "#395B74") +
  theme_minimal() +
  geom_text(data = filter(d, player == "Rasheed Wallace"),
    aes(x = college, y = player),
    label = "College", fontface = "bold",
    color = "#56a0d3",
    vjust = -1) +
  geom_text(data = filter(d, player == "Rasheed Wallace"),
    aes(x = nba, y = player),
    label = "NBA", fontface = "bold",
    color = "#395B74",
    vjust = -1) +
  labs(x = "Three-Point Attempt Rate (% of shots from 3)",
    y = "Player",
    title = "Three-Point Attempt Rates in Collegiate and Pro Careers", caption = "Source: UNC Record Book and Pro-Basketball Reference") +
  theme(title = element_text()) +
  scale_x_continuous(labels = scales::percent,
    breaks = c(0, 0.10, 0.20, 0.30, .40)) +
  coord_cartesian(xlim = c(0, 0.40)) +
  geom_text(aes(y = player, label = diff),
    x = 0.4, hjust = 0.25) +
  annotate(x = 0.4, y = "Rasheed Wallace", label = "Delta",
    geom = "text", vjust = -2,
    fontface = "bold",
    hjust = 1)
```
