---
title: "Sans Touchdown Histogram"
author: "Chris Gallo"
date: '2020-10-18'
slug: histogram-touchdown
tags:
- R Markdown
categories: R
type: post
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.width=12, fig.height=8)
```

```{r, echo=FALSE, message=FALSE}
library(tidyverse)
library(ggthemes)

data <- read.csv('no-tds.csv')

p1 <- ggplot(data, aes(x=Year)) + theme_economist() + scale_colour_economist() +
  geom_histogram(binwidth=1, color="black", fill="#FBFFB9") +
  theme(legend.position="bottom", legend.direction="horizontil",
        legend.title = element_blank()) +
  scale_x_continuous(breaks=seq(2000,2020,1)) +
  scale_y_continuous(breaks=seq(0,8,1)) +
  stat_bin(binwidth=1, geom="text", aes(label=..count..), vjust=-1.5) +
  labs(x="Year", y="Frequency", 
title = "NFL: How Many Teams Win Without a Touchdown Season-by-Season Since 2000",
        caption = "Data source: stathead.com")
p1
```

How many NFL teams have won a game without scoring a touchdown since 2000? 

This histogram answers that question. Since 2000, 68 total teams have won a game in the regular season or playoffs without scoring a touchdown. Only two of those games were in the playoffs. 

The data here is pulled from [stathead.com](https://stathead.com/) by Sports-Reference. A Stathead membership is like a library card to the largest sports history treasure trove. This data was pulled in less than two minutes.

This uses the ggplot package and economist theme.

Code is below . . .

```
library(tidyverse)
library(ggthemes)

data <- read.csv('no-tds.csv')

p1 <- ggplot(data, aes(x=Year)) + theme_economist() + scale_colour_economist() +
  geom_histogram(binwidth=1, color="black", fill="#FBFFB9") +
  theme(legend.position="bottom", legend.direction="horizontil",
        legend.title = element_blank()) +
  scale_x_continuous(breaks=seq(2000,2020,1)) +
  scale_y_continuous(breaks=seq(0,8,1)) +
  stat_bin(binwidth=1, geom="text", aes(label=..count..), vjust=-1.5) +
  labs(x="Year", y="Frequency", 
title = "NFL: How Many Teams Win Without a Touchdown Season-by-Season Since 2000",
        caption = "Data source: stathead.com")
p1
```
