---
title: "GT Tables"
author: "Chris Gallo"
date: '2020-09-19'
slug: gt-tables
tags:
- R Markdown
categories: R
type: post
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.width=12, fig.height=8)
```

```{r, echo=FALSE, message=FALSE}
library(dplyr)
library(gt)

d <- tibble::tribble(
 ~Season,  ~Type, ~Opponent, ~Differential,
 "2004", "NCAA 2nd Round", "Texas", "-3",
 "2005", "NCAA Championship", "Illinois", "+5",
 "2006", "NCAA 2nd Round", "George Mason", "-5",
 "2007", "NCAA Elite Eight", "Georgetown", "-12",
 "2008", "NCAA Semifinal", "Kanas", "-18", 
 "2009", "NCAA Championship", "Michigan State", "+17",
 "2010", "NIT Final", "Dayton", "-11",
 "2011", "NCAA Elite Eight", "Kentucky", "-7",
 "2012", "NCAA Elite Eight", "Kansas", "-13", 
 "2013", "NCAA 2nd Round", "Kansas", "-12",
 "2014", "NCAA 2nd Round", "Iowa State", "-2",
 "2015", "NCAA Sweet Sixteen", "Wisconsin", "-7",
 "2016", "NCAA Championship", "Vilianova", "-3",
 "2017", "NCAA Championship", "Gonzaga", "+6",
 "2018", "NCAA 2nd Round", "Texas A&M", "-21",
 "2019", "NCAA Sweet Sixteen", "Auburn", "-17",
 "2020", "ACC 2nd Round", "Syracuse", "-28"
)  %>%
  dplyr::arrange(desc(Season))

gt_tbl <- gt(d)

gt_tbl <- 
  gt_tbl %>%
  tab_header(
    title = "Carolina: Point Differential in Final Game of Each Season Since 2004",
    subtitle = "Tar Heels season has ended with a double-digit loss three seasons in a row."
  ) %>%
  tab_source_note(
    source_note = "Source: goheels.com"
  )

gt_tbl 

```

This is a table showing the point differential for the final game of each North Carolina men's basketball season since 2004. Carolina's season has ended with a double-digit loss three seasons in a row by a combined 66 points. 

This is my first use of the [gt package](https://github.com/rstudio/gt/). It's quite useuful to put together a quick table without many lines of code. The labeling feels much better than formattable or dt packages. 

Code is below . . .

```
library(dplyr)
library(gt)

d <- tibble::tribble(
 ~Season,  ~Type, ~Opponent, ~Differential,
 "2004", "NCAA 2nd Round", "Texas", "-3",
 "2005", "NCAA Championship", "Illinois", "+5",
 "2006", "NCAA 2nd Round", "George Mason", "-5",
 "2007", "NCAA Elite Eight", "Georgetown", "-12",
 "2008", "NCAA Semifinal", "Kanas", "-18", 
 "2009", "NCAA Championship", "Michigan State", "+17",
 "2010", "NIT Final", "Dayton", "-11",
 "2011", "NCAA Elite Eight", "Kentucky", "-7",
 "2012", "NCAA Elite Eight", "Kansas", "-13", 
 "2013", "NCAA 2nd Round", "Kansas", "-12",
 "2014", "NCAA 2nd Round", "Iowa State", "-2",
 "2015", "NCAA Sweet Sixteen", "Wisconsin", "-7",
 "2016", "NCAA Championship", "Vilianova", "-3",
 "2017", "NCAA Championship", "Gonzaga", "+6",
 "2018", "NCAA 2nd Round", "Texas A&M", "-21",
 "2019", "NCAA Sweet Sixteen", "Auburn", "-17",
 "2020", "ACC 2nd Round", "Syracuse", "-28"
)  %>%
  dplyr::arrange(desc(Season)) %>%

gt_tbl <- 
  gt_tbl %>%
  tab_header(
    title = "Carolina: Point Differential in Final Game of Each Season Since 2004",
    subtitle = "Tar Heels season has ended with a double-digit loss three seasons in a row."
  ) %>%
  tab_source_note(
    source_note = "Source: goheels.com"
  )
```
