---
title: "Scatter First Round"
author: "Chris Gallo"
date: '2020-08-21'
slug: tar-heel-nba-draft-picks
tags:
- R Markdown
- Fungo Friday
categories: R
type: post
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.width=12, fig.height=8)
```

```{r, echo=FALSE, message=FALSE}
library(ggplot2)
library(ggthemes)
library(ggrepel)

data <- read.csv("nba-picks.csv")


p1 <-ggplot(data, aes(Year, Pick, label = Player))+
  geom_text_repel() +
  scale_x_continuous(breaks=seq(1976,2020,4)) +
 scale_y_reverse(breaks =seq(0,30,5)) +
  labs(x="Year", y="Pick Number", 
title = "Carolina: NBA First Round Draft Picks Since 1976",
caption = "Data source: realgm.com") + theme_economist() + theme(plot.title=element_text(family="OfficinaSanITC-Book"),
        text=element_text(family="OfficinaSanITC-Book")) 
p1
```
---

Here is a scatter plot showing all NBA first round draft picks out of North Carolina since 1976. The draft has changed a ton. Fewer rounds. Expansion teams. Eligibility rules. 

But there is always round one. 

Random notes: 

- 16 of Carolina's 21 top five picks came from 1976 to 1998 
- [three instances of Tar Heels](https://twitter.com/dadgumboxscores/status/1296942976569151488?s=20) being drafted back-to-back in the top five
- the last Tar Heel picked in round one of the 2009 draft is the only one still in the league today

The data is pulled from [basketball.realgm.com](https://basketball.realgm.com/ncaa/conferences/Atlantic-Coast-Conference/1/North-Carolina/2/nba-draft). This plot uses the Economist theme and [ggrepel package](https://cran.r-project.org/web/packages/ggrepel/vignettes/ggrepel.html). 

Code is below . . .

```
library(ggplot2)
library(ggthemes)
library(ggrepel)

p1 <-ggplot(data, aes(Year, Pick, label = Player))+
  geom_text_repel() +
  scale_x_continuous(breaks=seq(1976,2020,4)) +
 scale_y_reverse(breaks =seq(0,30,5)) +
  labs(x="Year", y="Pick Number", 
title = "Carolina: NBA First Round Draft Picks Since 1976",
caption = "Data source: realgm.com") + theme_economist() + theme(plot.title=element_text(family="OfficinaSanITC-Book"),
        text=element_text(family="OfficinaSanITC-Book")) 
p1
```
