---
title: "Dot Drill"
author: "Chris Gallo"
date: '2020-08-28'
slug: dot-champs
tags:
- R Markdown
- Fungo Friday
categories: R
type: post
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.width=12, fig.height=8)
```

```{r, echo=FALSE, message=FALSE}
library(ggplot2)
library(ggthemes)

# define data
d <- tibble::tribble(
    ~sports, ~champs,
  "Soccer (Women)",  22, 
  "Field Hockey", 8, 
  "Basketball (Men)", 6, 
  "Lacrosse (Men)", 5,
  "Tennis (Women)", 4,
  "Soccer (Men)", 2,
  "Lacrosse (Women)", 2,
  "Basketball (Women)", 1,
  "Tennis (Men)", 1
  )


ggplot(d) +
    geom_point(aes(x = champs, y = reorder(sports, champs)), colour = "#16284c", size = 3) +  
    theme_economist() + theme(plot.title=element_text(family="OfficinaSanITC-Book"),
        text=element_text(family="OfficinaSanITC-Book")) +
    scale_x_continuous(breaks=seq(0,22,4)) +
    labs(x="Team National Championships", y="Sport", 
    title = "University of North Carolina: Team National Championships By Sport (includes NCAA, ITA)",
    caption = "Data source: goheels.com")
```
---

This is a dotchart showing the University of North Carolina team championships by sport. It includes NCAA and ITA (tennis) events for any teams. 

Women's soccer has 22 of Carolina's 51 team National Championships. While women's soccer owns 43 percent of all team titles, it hasn't won any of the last 10: 

- Women's Tennis (4)
- Field Hockey (2)
- Women's Lacrosse (2)
- Men's Basketball (1)
- Men's Tennis (1)

The data is pulled from [goheels.com](https://goheels.com/news/2015/6/10/205473190.aspx). This plot uses the Economist theme and leaned on [this tutorial](https://ggplot2.tidyverse.org/reference/geom_point.html). 

Code is below . . .

```
library(ggplot2)
library(ggthemes)

# define data
d <- tibble::tribble(
    ~sports, ~champs,
  "Soccer (Women)",  22, 
  "Field Hockey", 8, 
  "Basketball (Men)", 6, 
  "Lacrosse (Men)", 5,
  "Tennis (Women)", 4,
  "Soccer (Men)", 2,
  "Lacrosse (Women)", 2,
  "Basketball (Women)", 1,
  "Tennis (Men)", 1
  )


ggplot(d) +
    geom_point(aes(x = champs, y = reorder(sports, champs)), colour = "#16284c", size = 3) +  
    theme_economist() + theme(plot.title=element_text(family="OfficinaSanITC-Book"),
    text=element_text(family="OfficinaSanITC-Book")) +
    scale_x_continuous(breaks=seq(0,22,4)) +
    labs(x="NCAA Championships", y="Sport", 
    title = "University of North Carolina: Team National Championships By Sport (includes NCAA, ITA)",
    caption = "Data source: goheels.com")
```
