---
title: "Chicklet charts"
author: "Chris Gallo"
date: '2021-01-23'
slug: chicklet-charts
tags:
- R Markdown
categories: R
type: post
---

```{r setup, include=FALSE }
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, fig.width=12, fig.height=8)
```

```{r, echo=FALSE, message= FALSE, warning = FALSE}
library(tidyverse)
library(extrafont)
library(rvest)
library(janitor)
library(scales)
library(ggchicklet)
library(paletteer)

theme_me <- function () {
  theme_minimal(base_size = 15, base_family = "RobotoCondensed-Regular") %+replace%
    theme (
      plot.title = element_text(hjust = 0.5),
      plot.subtitle = element_text(
        hjust = 0.5,
        vjust = -2,
        lineheight = 0.9,
        size = 10
      ),
      plot.caption = element_text(size = 8, hjust = 1),
      panel.grid.minor = element_blank(),
      plot.background = element_rect(fill = "#c9e1f1", color = "#c9e1f1")
    )
}

df <- read.csv('ologs.csv')

df$type <- as.factor(df$type)
df$type <- factor(df$type, levels = c("paint", "non_paint_two", "three", "free_throws"))

df$game <- as.factor(df$game)
df$game <- factor(df$game, levels = c("NC State", "Georgia Tech", "Notre Dame", "Miami", "Syracuse", "Florida State", "Wake Forest"))


df %>%
  ggplot(aes(game, pct)) +
  geom_chicklet(aes(fill = type)) +
  geom_text(
    data = . %>% filter(
      pct > .1,
      game == "Wake Forest" | game == "Notre Dame" | game == "Miami" | game == "Georgia Tech" | game == "Florida State"
    ),
    aes(label = percent(pct, accuracy = 1L)),
    family = "RobotoCondensed-Regular",
    size = 4,
    position = position_stack(vjust = 0.15)
  ) +
  geom_text(
    data = . %>% filter(pct > .01, game == "Syracuse" | game == "NC State"),
    aes(label = percent(pct, accuracy = 1L)),
    family = "RobotoCondensed-Regular",
    size = 4,
    position = position_stack(vjust = 0.15)
  ) +
  scale_y_continuous(
    position = "left",
    labels = c("0%", "25%", "50%", "75%", "100%"),
    limits = c(0, 1)
  ) +
  theme_me() +
  coord_flip() +
  theme(
    legend.position = 'bottom',
    axis.title.y = element_blank(),
    axis.title.x = element_blank(),
    legend.title = element_blank(),
    legend.text = element_text(size = 9),
    plot.title = element_text(face = 'bold', hjust = .5),
    plot.subtitle = element_text(hjust = .5, size = 10),
    plot.title.position = "plot",
    plot.margin = unit(c(.5, .5, 1, .5), "lines"),
    axis.text.y = element_text(margin = margin(0, -15, 0, 0)),
    legend.margin = margin(0, 0, 0, 0),
    legend.box.margin = margin(.5, .5, .5, .5)
  ) +
  labs(title = "Carolina Offense: Points Diet in ACC Games as of January 22, 2021",
       subtitle = "Percentage of points from paint, non-paint two-point shots, three-point shots, and free throws.",
       caption = "@dadgumboxscores")  +
  guides(
    fill = guide_legend(
      keywidth = .5,
      keyheight = .2,
      default.unit = "inch",
      label.position = 'top',
      nrow = 1
    )
  ) + scale_fill_paletteer_d("nord::aurora",
                         labels = c("Paint", "Non-Paint Twos", "Threes", "Free Throws"))

```
This is plot showing points distribution for Carolina basketball across two-pointers in the paint, two-pointers outside the paint, three-pointers, and free throws. 

It uses the [ggchicklet](https://github.com/hrbrmstr/ggchicklet) package, and it was inspired by Owen J. Phillips' awesome [newsletter - F5](https://thef5.substack.com/).

This one was particularly tricky to apply the text of the percentages. A few required different adjusting to show the percentages on the graph.  

Code is below . . .

```
library(tidyverse)
library(extrafont)
library(rvest)
library(janitor)
library(scales)
library(ggchicklet)
library(paletteer)

theme_me <- function () {
  theme_minimal(base_size = 15, base_family = "RobotoCondensed-Regular") %+replace%
    theme (
      plot.title = element_text(hjust = 0.5),
      plot.subtitle = element_text(
        hjust = 0.5,
        vjust = -2,
        lineheight = 0.9,
        size = 10
      ),
      plot.caption = element_text(size = 8, hjust = 1),
      panel.grid.minor = element_blank(),
      plot.background = element_rect(fill = "#c9e1f1", color = "#c9e1f1")
    )
}

df <- read.csv('unclogs.csv')

df$type <- as.factor(df$type)
df$type <-
  factor(df$type,
         levels = c("paint", "non_paint_two", "three", "free_throws"))

df$game <- as.factor(df$game)
df$game <-
  factor(
    df$game,
    levels = c(
      "NC State",
      "Georgia Tech",
      "Notre Dame",
      "Miami",
      "Syracuse",
      "Florida State",
      "Wake Forest"
    )
  )


df %>%
  ggplot(aes(game, pct)) +
  geom_chicklet(aes(fill = type)) +
  geom_text(
    data = . %>% filter(
      pct > .1,
      game == "Wake Forest" |
        game == "Notre Dame" |
        game == "Miami" |
        game == "Georgia Tech" |
        game == "Florida State"
    ),
    aes(label = percent(pct, accuracy = 1L)),
    family = "RobotoCondensed-Regular",
    size = 4,
    position = position_stack(vjust = 0.15)
  ) +
  geom_text(
    data = . %>% filter(pct > .01, game == "Syracuse" |
                          game == "NC State"),
    aes(label = percent(pct, accuracy = 1L)),
    family = "RobotoCondensed-Regular",
    size = 4,
    position = position_stack(vjust = 0.15)
  ) +
  scale_y_continuous(
    position = "left",
    labels = c("0%", "25%", "50%", "75%", "100%"),
    limits = c(0, 1)
  ) +
  theme_me() +
  coord_flip() +
  theme(
    legend.position = 'bottom',
    axis.title.y = element_blank(),
    axis.title.x = element_blank(),
    legend.title = element_blank(),
    legend.text = element_text(size = 9),
    plot.title = element_text(face = 'bold', hjust = .5),
    plot.subtitle = element_text(hjust = .5, size = 10),
    plot.title.position = "plot",
    plot.margin = unit(c(.5, .5, 1, .5), "lines"),
    axis.text.y = element_text(margin = margin(0, -15, 0, 0)),
    legend.margin = margin(0, 0, 0, 0),
    legend.box.margin = margin(.5, .5, .5, .5)
  ) +
  labs(title = "Carolina Offense: Points Diet in ACC Games as of January 22, 2021",
       subtitle = "Percentage of points from paint, non-paint two-point shots, three-point shots, and free throws.",
       caption = "@dadgumboxscores")  +
  guides(
    fill = guide_legend(
      keywidth = .5,
      keyheight = .2,
      default.unit = "inch",
      label.position = 'top',
      nrow = 1
    )
  ) +
  scale_fill_paletteer_d("nord::aurora",
                         labels = c("Paint", "Non-Paint Twos", "Threes", "Free Throws"))
```
